\documentclass[a4paper]{article}
\usepackage[fleqn]{amsmath}
\usepackage{graphicx}
%\usepackage{times}
\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
\usepackage{listing}
\usepackage[small,compact]{titlesec}
\usepackage[utf8]{inputenc}

\usepackage{biblatex}
\bibliography{../Documentation}

\usepackage[paper=a4paper,
            includefoot, % Uncomment to put page number above margin
            marginparwidth=30.5mm,    % Length of section titles
            marginparsep=1.5mm,       % Space between titles and text
            margin=10mm,              % 25mm margins
            includemp]{geometry}

%\setlength{\oddsidemargin}{10mm}
%\setlength{\evensidemargin}{10mm}
\usepackage{fullpage}

\usepackage{multicol}
\usepackage{caption}

\newcommand{\makeheading}[2]%
        {\hspace*{-\marginparsep minus \marginparwidth}%
         \begin{minipage}[t]{\textwidth\marginparwidth\marginparsep}%
           {\large \bfseries #1}\\{#2}\\[-0.15\baselineskip]%
                 \rule{\columnwidth}{1pt}%
         \end{minipage}}

\newlength{\figurewidth}
\setlength{\figurewidth}{500px}


\begin{document}
\makeheading{Gautebøye - Processing}{Gaute Hope
(gaute.hope@student.uib.no), 21.08.2012, Revision 1}

\vspace{2em}
\section*{Introduction}
This document describes the simple processing tools as well as some
notes for how to use the data with SEISAN or MATLAB.

\vspace{2em}

\begin{multicols}{2}
  \section{Tools}
  \subsection{dtttomseed}
  \subsection{dattodtt}
  \subsection{dttfix}
  \subsection{fakedtt}
  \subsection{mschangesource}

  \section{SEISAN}
    \subsection{Configuration}
    You have to increase the plot resolution of mulplt (MULPLT.DEF) to something like
    60000 (screen resolution and hardcopy resolution), as well as update the
    frequency band of the spectral analysis to at least 125 Hz. More
    might be required. There is 40960 points in each DAT or DTT file.

    \subsection{Conversion}
    Use \textit{dtttomseed} or \textit{dattomseed} to convert to miniSEED, update the source
    file to change location, station is given as first argument. I have used SHZ as channel code
    for the hydrophone.

    \subsection{Test}
    Use the program \textit{fakedtt.py} to create a DTT file with a sine
    wave and try to convert and plot it to see if everything works.

    \subsection{Importing to SEISAN}
      \begin{enumerate}
        \item Create a work directory, copy the data and index files for
          the events to it.
        \item Use dtttomseed or dattomseed with the first argument being
          station name, BUO1 to BUO5 and after that a list of sequences
          of the ids of the files.
        \item A mseed file is created in the current directory.
        \item Use mscut to split the file up in hourly intervals.
        \item Create a list of the final files with dirf.
        \item Use autoreg to add the events to the database. Copy the
          WAV files and select create new IDs if they conflict with
          existing events.
        \item Use associ to merge the events with the existing events in
          the database. Delete merged events.
      \end{enumerate}

      \subsection{Continuous database setup}
      \begin{enumerate}
        \item Set up a database for the experiment, e.g. GAK1
        \item Import events hourly as described above, one event for
          each hour.
        \item Set up SEISAN.DEF to know GAK1 as a continuous database.
        \item Set a new database for registered and extracted events, e.g. GAK1E.
        \item Work on GAK1 and register ('p' in mulplt) to create an
          event in GAK1E when something interesting is found.
        \item GAK1 can be used for noise measurements etc.
        \item Possibly create an intermittent stop with sanitized events
          (fixed time, etc..)
      \end{enumerate}
  \section{MATLAB}

%\vspace{1em}
%\printbibliography

\end{multicols}
\end{document}

